Challenges to the Two-Infall Scenario by Large Stellar Age Catalogs

### 2025-07-31 src/tex/ms.tex
Manuscript updates

\begin{abstract}
    Stars in the Milky Way disk exhibit a clear separation into two chemically distinct populations based on their [$\alpha$/Fe] ratios. This $\alpha$-bimodality is not a universal feature of simulated disk galaxies and may point to a unique evolutionary history. A popular explanation is the two-infall scenario, which postulates that two periods of substantial accretion rates dominate the assembly history of the Galaxy. 
    % However, most previous studies using the two-infall scenario have explored a limited portion of the parameter space, typically neglecting radial migration and assuming that the Galactic disk never ejected a substantial outflow. 
    Thanks to recent advances in stellar age measurements, we can now compare this popular model to more direct measurements of the Galaxy's evolutionary timescales across the disk. 
    % We run multi-zone galactic chemical evolution models with a two-infall-driven star formation history, radially dependent mass-loaded outflows, and a prescription for radial migration tuned to a hydrodynamical simulation. We compare our model results to abundance patterns across the disk from APOGEE DR17, supplemented with stellar age estimates through multiple methods. 
    We run multi-zone galactic chemical evolution models with a two-infall-driven star formation history and compare the results against abundance patterns across the disk from APOGEE DR17, supplemented with stellar ages estimated through multiple methods.
    Although the two-infall scenario offers a natural explanation for the $\alpha$-bimodality, it struggles to explain several features of the age--abundance structure in the disk. First, our models generically predict a massive and long-lasting dilution event, but the data show that stellar metallicity is remarkably constant across much of the lifetime of the disk. This apparent age-independence places considerable restrictions upon the two-infall parameter space. Second, most local metal-rich stars in APOGEE have intermediate ages, yet our models predict these stars should either be very old or very young. Some of these issues can be mitigated, but not completely resolved, by allowing the accreted gas to be pre-enriched to low metallicity. 
    % Restrictions upon the two-infall parameter space also limit the application of other merger-dominated star formation histories to the Milky Way.
    These restrictions also place limits on the role of merger events in shaping the chemical evolution of the thin disk in general.
\end{abstract}

@@

% \todo{Hook, summary of alpha-bimodality.}
\lettrine{G}{alactic chemical evolution} (GCE) aims to explain the metal abundance patterns observed in the Milky Way (MW) by modeling the star formation history and evolution of the Galaxy. A long-standing paradigm of GCE is that the metallicity of the interstellar medium (ISM) increases over time due to supernova enrichment from successive generations of stars \citep[e.g.,][]{tinsley_stellar_1979,matteucci_relative_1986}. In this view, one feature of the MW disk that is difficult to explain is the so-called ``$\alpha$-bimodality'': the presence of two populations of stars at similar metallicity but separated by their \aFe ratio \citep[e.g.,][]{bensby_exploring_2014}. The high-$\alpha$ sequence consists of old stars \citep[$\gtrsim9$ Gyr; e.g.,][]{pinsonneault_apokasc-3_2025} with super-Solar \aFe and is associated with the kinematic thick disk \citep[e.g.,][]{fuhrmann_nearby_1998}, while the low-$\alpha$ sequence is younger, with approximately Solar \aFe, and is associated with the kinematic thin disk. The $\alpha$-bimodality is present across the Galactic disk, but the relative strength of the high- and low-$\alpha$ sequences varies by location \citep{hayden_chemical_2015}.

An explanation for the MW $\alpha$-bimodality has yet to be broadly accepted in the GCE literature. An $\alpha$-bimodality is not a universal feature in simulated MW-mass galaxies \citep[e.g.,][]{mackereth_origin_2018,parul_effect_2025}, and tentatively does not exist in M31 (\citealt{nidever_prevalence_2024}; but see also \citealt{kobayashi_fe_2023}). Its presence and characteristics in our Galaxy could potentially have arisen to a unique evolutionary history. GCE models that attempt to solve this problem generally fall into two camps. Some explain the $\alpha$-bimodality as a result of secular processes, such as the radial migration of stars and the inside-out growth of the disk \citep[e.g.,][]{schonrich_chemical_2009,kubryk_evolution_2015,sharma_chemical_2021,chen_chemical_2023,prantzos_origin_2023}. Others argue for a bursty star formation history driven by multiple accretion events \citep[e.g.,][]{chiappini_chemical_1997,mackereth_origin_2018,spitoni_beyond_2023}.
% Studies of simulations have suggested that the $\alpha$-bimodality could be driven by clumpy star formation at high redshift \citep{clarke_imprint_2019}, a major gas-rich merger \citep{buck_origin_2020}, or re-accretion of metal-rich outflows \citep{khoperskov_bimodality_2021}.

@@

By neglecting Galactic outflows, previous studies of the two-infall scenario have been constrained in their choice of nucleosynthetic yields \citep{francois_evolution_2004} because of the yield--outflow degeneracy \citep[e.g.,][]{hartwick_chemical_1976,cooke_primordial_2022,johnson_dwarf_2023,sandford_strong_2024}. \citet{weinberg_equilibrium_2017} showed that the equilibrium metallicity is primarily set by yields and the outflow mass-loading factor; proportionally raising or lowering both may affect the time-dependence of abundance evolution but not the end-point. This degeneracy prohibits direct estimates of the yield scale from GCE models, unless the effect of outflows is assumed to be insignificant. The predicted yields from CCSN models can vary substantially depending on the choice of initial mass function \citep{vincenzo_modern_2016} and the physics of black hole formation \citep{griffith_impact_2021}, yet few studies have investigated the effect of the yield scale on two-infall scenario predictions. Varying the yield scale while maintaining an evolutionary end point that is consistent with observations is straightforwardly achieved with simultaneous variations in the strength of outflows.

@@

% \todo{Migration.} 
Previous two-infall models have reproduced the full distribution of stellar abundances in the Solar neighborhood through a single, continuous evolutionary track. However, the current body of evidence suggests that many of the stars that make up the wings of the local metallicity distribution originate from elsewhere in the Galaxy. \citet{sellwood_radial_2002} first showed that transient spiral perturbations can induce large changes in the guiding radius of a star without kinematic heating, and it is now understood that the stars that make up the Solar neighborhood are drawn from a wide range of birth radii \citep[e.g.,][]{schonrich_chemical_2009,frankel_measuring_2018,lehmann_probing_2024}. Some studies have attempted to derive stellar birth radii \citep[e.g.,][]{ratcliffe_unveiling_2023,lu_there_2024}, though such an endeavor depends on a model of the evolution of the radial metallicity profile in the MW. While the strength and speed of radial migration in the disk are not precisely measured, a single chemical evolution track need not explain the entire width of the local abundance distribution. We illustrate this distinction in Figure \ref{fig:smooth-vs-twoinfall}, which contrasts the chemical evolution track for the Solar neighborhood predicted by a two-infall model against tracks from the smooth SFH model of \citet{johnson_stellar_2021} at three different radii.

% \todo{Summary of secular evolution models which claim alpha-bimodality.} 
In contrast to the two-infall scenario, a number of studies have attempted to explain the $\alpha$-bimodality through purely secular processes. 
\citet{schonrich_chemical_2009} first showed that radial migration can produce an $\alpha$-bimodality through a superposition of chemical tracks, although they did not directly compare against MW data. Others have produced a more MW-like $\alpha$-bimdodality using a combination of radial migration and inside-out galaxy growth \citep[e.g.,][]{kubryk_evolution_2015,sharma_chemical_2021,chen_chemical_2023,prantzos_origin_2023}. In this scenario, the local high-$\alpha$ population originates from the inner Galaxy, where the star formation rate peaked early in its history. \citet{sharma_chemical_2021} and \citet{chen_chemical_2023} showed that a simultaneous decline in the star formation rate and the \aFe ratio is needed to separate the sequences in chemical space under this scenario. \citet{chen_recent_2025} additionally argued that the double sequence in the local age--metallicity relation observed by \citet{nissen_high-precision_2020} can also be explained by a smooth star formation history and inside-out disk growth. However, other GCE models that incorporate both radial migration and smooth, inside-out star formation have failed to predict an $\alpha$-bimodality \citep[e.g.,][]{johnson_stellar_2021,dubay_galactic_2024}. While the $\alpha$-bimodality remains a key piece of evidence for the two-infall scenario, it is not the only proposed explanation.

@@

% \todo{Introduce equilibrium model.} 
As stellar age estimation techniques have improved over recent years, large catalogs have become available with measurements for large samples of stars from a wide swath of the Galaxy. In a challenge to the traditional view of GCE, which expects the ISM metallicity to continually increase over time, 
% \citet{johnson_milky_2024} examined the age--metallicity relation at different radii from the astroNN catalog \citep{mackereth_dynamical_2019} and found that the mode of the metallicity distribution at a given radius is nearly independent from age over the past $\sim9\Gyr$. They propose an
these catalogs have indicated that stellar metallicities are remarkably consistent with age across most of the thin disk epoch \citep[e.g.,][]{spina_mapping_2022,magrini_gaia-eso_2023,willett_evolution_2023,carbajo-hijarrubia_occaso_2024,gallart_chronology_2024}. \citet{johnson_milky_2024} demonstrated that these results are readily explained by a scenario in which the Galaxy reaches a state of chemical equilibrium. In this so-called ``equilibrium scenario,'' the local metallicity is driven by the ratio of star formation to accretion at a given radius, which evolves to a constant over Gyr timescales. Whether the equilibrium metallicity is regulated by outflows, as in the  \citet{johnson_milky_2024} models, or by other factors such as a radial gas flow \citep{spitoni_effects_2011,bilitewski_radial_2012,sharda_interplay_2024}, the current data suggest that metal abundances in the ISM evolved minimally during the thin disk epoch.

% In light of the findings of \citet{johnson_milky_2024} and a new empirical yield scale from \citet{weinberg_scale_2024}, we evaluate the predictions of the two-infall model against stellar age and abundance data across the MW disk.
Relative to even a few years ago, the availability of large stellar catalogs with [Fe/H], \aFe, {\it and} age covering a large portion of the Galactic disk enables much more comprehensive tests of the two-infall GCE scenario (and other models). Our study draws on the steadiness of the Milky Way stellar metallicity gradient found by \citet{johnson_milky_2024} and empirical constraints on the yield scale from \citet{weinberg_scale_2024}.
% \todo{Goal and summary of paper.} 
We run multi-zone GCE models with a two-infall accretion history, radially-dependent mass-loaded outflows, and a prescription for radial migration tuned to a hydrodynamical simulation. We investigate the impact of the scale of SN yields and outflows, the strength of radial migration, the composition of the circumgalactic medium, and the local disk mass surface density ratio on GCE model predictions. We compare our results to abundance distributions across the disk from APOGEE DR17, and to age--abundance relations from multiple age catalogs. We describe our observational sample in Section \ref{sec:observational-sample}, and we detail our chemical evolution models and parameter selection in Section \ref{sec:methods}. We compare our multi-zone model predictions to the data in Section \ref{sec:multizone-results}. We discuss our results in Section \ref{sec:discussion} and summarize our conclusions in Section \ref{sec:conclusions}.

@@

We obtain a sample of \num{171635} red giant branch and red clump stars with high-quality spectra using the selection criteria listed in Table \ref{tab:sample}, which are adapted from \citet{hayden_chemical_2015}. Table \ref{tab:uncertainties} presents the median statistical uncertainty and uncertainty dispersion ($95^{\rm th} - 5^{\rm th}$ percentile difference) of the calibrated [Fe/H] and [O/Fe] abundances for our sample. When calculating the galactocentric radius $R_{\rm gal}$ and midplane distance $z$ of each star, we use the \citet{bailer-jones_estimating_2021} photo-geometric distance estimates from {\it Gaia} Early Data Release 3 \citep{gaia_collaboration_gaia_2016,gaia_collaboration_gaia_2021} included in the APOGEE DR17 catalog and we adopt the Galactic coordinates of the Sun $(R,z)_\odot=(8.122, 0.0208)\kpc$ \citep{gravity_collaboration_detection_2018,bennett_vertical_2019}. 
% Table \ref{tab:apogee-regions} presents the number of APOGEE stars in bins of $R_{\rm gal}$ and $|z|$; typical distance uncertainties are much smaller than the bin width \todo{(do I need this table?)}.

\subsection{Stellar Age Estimates}

@@

Our second age catalog utilizes the [C/N]--age relation calibrated by Roberts et al.\ (2025, submitted) for red giant branch (RGB) and red clump stars. The relationship relies on the mass-dependent level of mixing during first dredge-up \citep[FDU;][]{iben_stellar_1967} to map the correlation of stellar mass, and hence age, with surface chemistry. This method has the benefit of providing age estimates for luminous giants ($\log g<2.5$), which increases the sample size at larger distances from the Sun. Limitations from the efficiency of FDU mixing and the RGB age--mass relationship mean the ages are not trustworthy outside the range $1\sim10\Gyr$, but this is the age range most useful for our purposes anyway. Additional mixing effects in low-metallicity stars %later in their RGB evolution 
also prevent the relation from being applied to luminous giant and red clump stars with $\mathFeH<-0.4$; stars on the lower RGB do not suffer from this problem. The median propagated uncertainty for the [C/N]-derived ages is $\sim1\Gyr$; however, as noted by Roberts et al.\ (2025, submitted), the propagated errors underestimate the true age uncertainty, so we enhance the uncertainties by 40\% (see Table \ref{tab:uncertainties}). With this relationship, we estimate ages for \num{113464} stars across the disk, including \num{20995} in the Solar neighborhood.

@@

% We assign a final radius and midplane distance to each stellar population as described in Section \ref{sec:migration}. 
We discuss our assumptions about the nucleosynthetic yields in Section \ref{sec:yields}, the outflow prescription in Section \ref{sec:outflows}, the star formation law in Section \ref{sec:sf-law}, the gas supply equation in Section \ref{sec:sfh}, the infall parameter selection in Section \ref{sec:parameter-selection}, and the stellar migration prescription in Section \ref{sec:migration}. We do not incorporate radial gas flows between the different zones, but we discuss their potential implications in Section \ref{sec:radial-flows}. Table \ref{tab:parameters} summarizes the most relevant variables and their fiducial values in this work.

@@

        M_{\rm \star,tot}   & $5.17\times 10^{10}\Msun$     & ---   & \ref{sec:sfh} & Total stellar mass of the disk \\

@@

\begin{table}
    \centering
    \caption{Nucleosynthetic yields and outflow prescriptions (see Section \ref{sec:yields}).}
    \include{output/yields.tex}
    \label{tab:yields}
\end{table}

@@

Unlike CCSNe, SNe Ia populate a broad distribution of delay times between progenitor formation and explosion. The time-dependent SN Ia rate per unit mass of star formation is defined as
\begin{equation}
    R_{\rm Ia}(t) = 
    \begin{cases}
        \frac{N_{\rm Ia}}{M_\star}
        \frac{f_{\rm Ia}(t)}{\int_{t_{\rm Ia}}^{t_{\rm max}} f_{\rm Ia}(t') dt'}, & t \ge t_{\rm Ia} \\
        0 & t < t_{\rm Ia},
    \end{cases}
    \label{eq:dtd-function}
\end{equation}
where $t_{\rm Ia}=40\,{\rm Myr}$ is the minimum SN Ia delay time, $t_{\rm final}=13.2\,{\rm Gyr}$ is our assumed lifetime of the Galactic disk, and $f_{\rm Ia}(t)$ is the un-normalized form of the DTD. Motivated by recent results suggesting that a large fraction of long-delayed SNe Ia improves agreement with the Milky Way's high-$\alpha$ sequence \citep{palicio_analytic_2023,dubay_galactic_2024}, we adopt a wide plateau DTD of the form
\begin{equation}
    \label{eq:plateau-dtd}
    f_{\rm Ia}(t) =
    \begin{cases}
        1, & t < 1\,{\rm Gyr} \\
        (t/1\,{\rm Gyr})^{-1.1}, & t \ge 1\,{\rm Gyr}.
    \end{cases}
\end{equation}
We also explore a simple power-law DTD,
\begin{equation}
    f_{\rm Ia}^{\rm plaw}(t) = (t/1\,\rm{Gyr})^{-1.1}.
    \label{eq:powerlaw-dtd}
\end{equation}
Equation \ref{eq:powerlaw-dtd} is consistent with SN Ia surveys \citep[e.g.,][]{maoz_star_2017} and results in more prompt Fe production.
We discuss the implications of using a different DTD in Section \ref{sec:abundance-distributions}.

@@

Mass-loaded outflows are a useful tool for scaling the endpoint of GCE models. \citet{weinberg_equilibrium_2017} showed that in the case of exponentially declining star formation, the O abundance approaches an equilibrium at
\begin{equation}
    \label{eq:equilibrium}
    Z_{\rm O,eq} = \frac{y_{\rm O}^{\rm CC}}{1 + \eta - r - \tau_\star/\tau_{\rm SFH}},
\end{equation}
where $r=0.4$ is the instantaneous recycling parameter, $\tau_\star$ is the star formation efficiency timescale, $\tau_{\rm SFH}$ is the $e$-folding timescale of the star formation history, and $\eta\equiv \dot\Sigma_{\rm out}/\dot\Sigma_\star$ is the outflow mass-loading factor. Motivated by Equation \ref{eq:equilibrium}, we adopt the outflow mass-loading factor at the Solar radius $\eta_\odot\equiv\eta(R=R_\odot)$ for each of the yield sets in Table \ref{tab:yields}. Models with lower yields do not achieve a steady-state abundance in time (see Figure \ref{fig:yield-outflow}); therefore, the values of $\eta_\odot$ for \yZ{1} and \yZ{2} are lower than Equation \ref{eq:equilibrium} would indicate as required to reach Solar metallicity at the present day, assuming that $Z_{\rm O,eq}\approx Z_{\rm O,\infty}$ as suggested by \citet{johnson_milky_2024} based on age trends such as these.

@@

We note that all our models are normalized to produce the same thick-to-thin-disk mass ratio of $f_{\Sigma}(R_\odot)=0.12$ \citep{bland-hawthorn_galaxy_2016} at the Solar annulus regardless of the infall parameters. The high-$\alpha$ sequence appears much less prominent in our [O/Fe] distributions in Figure \ref{fig:twoinfall-parameters} than in the data because the model outputs include only stars which were formed in-situ at the Solar annulus, since they are one-zone models. In our multi-zone models, most of the high-$\alpha$ stars present in the Solar neighborhood have migrated from the inner Galaxy.

>>>>

\begin{abstract}
    Stars in the Milky Way disk exhibit a clear separation into two chemically distinct populations based on their [$\alpha$/Fe] ratios. This $\alpha$-bimodality is not a universal feature of simulated disk galaxies and may point to a unique evolutionary history. A popular explanation is the two-infall scenario, which postulates that two periods of substantial accretion rates dominate the assembly history of the Galaxy. Thanks to recent advances in stellar age measurements, we can now compare this popular model to more direct measurements of the Galaxy's evolutionary timescales across the disk. We run multi-zone galactic chemical evolution models with a two-infall-driven star formation history and compare the results against abundance patterns across the disk from APOGEE DR17, supplemented with stellar ages estimated through multiple methods. Although the two-infall scenario offers a natural explanation for the $\alpha$-bimodality, it struggles to explain several features of the age--abundance structure in the disk. First, our models generically predict a massive and long-lasting dilution event, but the data show that stellar metallicity is remarkably constant across much of the lifetime of the disk. This apparent age-independence places considerable restrictions upon the two-infall parameter space. Second, most local metal-rich stars in APOGEE have intermediate ages, yet our models predict these stars should either be very old or very young. Some of these issues can be mitigated, but not completely resolved, by pre-enriching the accreted gas to low metallicity. These restrictions also place limits on the role of merger events in shaping the chemical evolution of the thin disk in general.
\end{abstract}

@@

% \todo{Hook, summary of alpha-bimodality.}
\lettrine{G}{alactic chemical evolution} (GCE) aims to explain the metal abundance patterns observed in the Milky Way (MW) by modeling the star formation history and evolution of the Galaxy. A long-standing paradigm of GCE is that the metallicity of the interstellar medium (ISM) increases over time due to supernova enrichment from successive generations of stars \citep[e.g.,][]{tinsley_stellar_1979,matteucci_relative_1986}. In this view, one feature of the MW disk that is difficult to explain is the so-called ``$\alpha$-bimodality'': the presence of two populations of stars at similar metallicity but separated by their \aFe ratio \citep[e.g.,][]{bensby_exploring_2014}. The high-$\alpha$ sequence consists of old stars \citep[$\gtrsim9$ Gyr; e.g.,][]{pinsonneault_apokasc-3_2025} with super-Solar \aFe and is associated with the kinematic thick disk \citep[e.g.,][]{fuhrmann_nearby_1998}, while the low-$\alpha$ sequence is younger, has approximately Solar \aFe, and is associated with the kinematic thin disk. The $\alpha$-bimodality is present across the Galactic disk, but the relative strength of the high- and low-$\alpha$ sequences varies by location \citep{hayden_chemical_2015}.

An explanation for the MW $\alpha$-bimodality has yet to be broadly accepted in the GCE literature. An $\alpha$-bimodality is not a universal feature in simulated MW-mass galaxies \citep[e.g.,][]{mackereth_origin_2018,parul_effect_2025}, and tentatively does not exist in M31 (\citealt{nidever_prevalence_2024}; but see also \citealt{kobayashi_fe_2023}). Its presence and characteristics in our Galaxy could potentially have arisen due to a unique evolutionary history. GCE models that attempt to solve this problem generally fall into two camps. Some explain the $\alpha$-bimodality as a result of secular processes, such as the radial migration of stars and the inside-out growth of the disk \citep[e.g.,][]{schonrich_chemical_2009,kubryk_evolution_2015,sharma_chemical_2021,chen_chemical_2023,prantzos_origin_2023}. Others argue for a bursty star formation history driven by multiple accretion events \citep[e.g.,][]{chiappini_chemical_1997,mackereth_origin_2018,spitoni_beyond_2023}.
% Studies of simulations have suggested that the $\alpha$-bimodality could be driven by clumpy star formation at high redshift \citep{clarke_imprint_2019}, a major gas-rich merger \citep{buck_origin_2020}, or re-accretion of metal-rich outflows \citep{khoperskov_bimodality_2021}.

@@

By neglecting Galactic outflows, previous studies of the two-infall scenario have been constrained in their choice of nucleosynthetic yields \citep{francois_evolution_2004} because of the yield--outflow degeneracy \citep[e.g.,][]{hartwick_chemical_1976,cooke_primordial_2022,johnson_dwarf_2023,sandford_strong_2024}. \citet{weinberg_equilibrium_2017} showed that the equilibrium metallicity is primarily set by the yields and the outflow mass-loading factor; proportionally raising or lowering both may affect the time-dependence of abundance evolution but not the end-point. This degeneracy prohibits direct estimates of the yield scale from GCE models, unless the effect of outflows is assumed to be insignificant. The predicted yields from CCSN models can vary substantially depending on the choice of initial mass function \citep{vincenzo_modern_2016} and the physics of black hole formation \citep{griffith_impact_2021}, yet few studies have investigated the effect of the yield scale on two-infall scenario predictions. Varying the yield scale while maintaining an evolutionary end point that is consistent with observations is straightforwardly achieved with simultaneous variations in the strength of outflows.

@@

% \todo{Migration.} 
Previous two-infall models have reproduced the full distribution of stellar abundances in the Solar neighborhood through a single, continuous evolutionary track. However, the current body of evidence suggests that many of the stars that make up the wings of the local metallicity distribution originate from elsewhere in the Galaxy. \citet{sellwood_radial_2002} first showed that transient spiral perturbations can induce large changes in the guiding radius of a star without kinematic heating, and it is now understood that the stars that make up the Solar neighborhood are drawn from a wide range of birth radii \citep[e.g.,][]{schonrich_chemical_2009,frankel_measuring_2018,lehmann_probing_2024}. Some studies have attempted to derive stellar birth radii \citep[e.g.,][]{ratcliffe_unveiling_2023,lu_there_2024}, though such an endeavor depends on a model of the evolution of the radial metallicity profile in the MW. While the strength and speed of radial migration in the disk are not precisely measured, a single chemical evolution track need not explain the entire width of the local abundance distribution. We illustrate this distinction in Figure \ref{fig:smooth-vs-twoinfall}, which contrasts the chemical evolution track for the Solar neighborhood predicted by a two-infall model against tracks from the smooth SFH model of \citet{johnson_stellar_2021} at three different radii. If the effect of radial migration is significant, then the full local stellar abundance distribution can be explained without dramatic changes to the local gas abundance over time.

% \todo{Summary of secular evolution models which claim alpha-bimodality.} 
In contrast to the two-infall scenario, a number of studies have attempted to explain the $\alpha$-bimodality through purely secular processes. 
\citet{schonrich_chemical_2009} first showed that radial migration can produce an $\alpha$-bimodality through a superposition of chemical tracks, although they did not directly compare against MW data. Others have produced a more MW-like $\alpha$-bimdodality using a combination of radial migration and inside-out galaxy growth \citep[e.g.,][]{kubryk_evolution_2015,sharma_chemical_2021,chen_chemical_2023,prantzos_origin_2023}. In this scenario, the local high-$\alpha$ population originates from the inner Galaxy, where the star formation rate peaked early in its history. \citet{sharma_chemical_2021} and \citet{chen_chemical_2023} showed that a simultaneous decline in the star formation rate and the \aFe ratio is needed to separate the sequences in chemical space. \citet{chen_recent_2025} additionally argued that the double sequence in the local age--metallicity relation observed by \citet{nissen_high-precision_2020} can also be explained by a smooth star formation history and inside-out disk growth. However, other GCE models that incorporate both radial migration and smooth, inside-out star formation have failed to predict an $\alpha$-bimodality \citep[e.g.,][]{johnson_stellar_2021,dubay_galactic_2024}. While the $\alpha$-bimodality remains a key piece of evidence for the two-infall scenario, it is not the only proposed explanation.

@@

% \todo{Introduce equilibrium model.} 
As stellar age estimation techniques have improved over recent years, large catalogs have become available with measurements for stars across a wide swath of the Galaxy. In a challenge to the traditional view of GCE, which expects the ISM metallicity to continually increase over time, 
% \citet{johnson_milky_2024} examined the age--metallicity relation at different radii from the astroNN catalog \citep{mackereth_dynamical_2019} and found that the mode of the metallicity distribution at a given radius is nearly independent from age over the past $\sim9\Gyr$. They propose an
these catalogs have indicated that stellar metallicities are remarkably consistent with age across most of the thin disk epoch \citep[e.g.,][]{spina_mapping_2022,magrini_gaia-eso_2023,willett_evolution_2023,carbajo-hijarrubia_occaso_2024,gallart_chronology_2024}. \citet{johnson_milky_2024} demonstrated that these results are readily explained by a scenario in which the Galaxy reaches a state of chemical equilibrium. In this so-called ``equilibrium scenario,'' the local metallicity is driven by the ratio of star formation to accretion at a given radius, which evolves to a constant over Gyr timescales. Whether the equilibrium metallicity is regulated by outflows, as in the  \citet{johnson_milky_2024} models, or by other factors such as a radial gas flow \citep{spitoni_effects_2011,bilitewski_radial_2012,sharda_interplay_2024}, the current data suggest that metal abundances in the ISM evolved minimally during the thin disk epoch.

% In light of the findings of \citet{johnson_milky_2024} and a new empirical yield scale from \citet{weinberg_scale_2024}, we evaluate the predictions of the two-infall model against stellar age and abundance data across the MW disk.
Relative to even a few years ago, the availability of large stellar catalogs with [Fe/H], \aFe, {\it and} age covering a large portion of the Galactic disk enables much more comprehensive tests of the two-infall GCE scenario (and other models). Our study draws on the steadiness of the Milky Way stellar metallicity gradient found by \citet{johnson_milky_2024} and empirical constraints on the yield scale from \citet{weinberg_scale_2024}.
% \todo{Goal and summary of paper.} 
We run multi-zone GCE models with a two-infall accretion history, radially-dependent mass-loaded outflows, and a prescription for radial migration tuned to a hydrodynamical simulation. We investigate the impact of the scale of SN yields and outflows, the strength of radial migration, the composition of the circumgalactic medium, and the local disk mass surface density ratio on GCE model predictions. We compare our results to abundance distributions across the disk from APOGEE DR17, and to age--abundance relations from multiple age catalogs. We describe our observational sample in Section \ref{sec:observational-sample}, and we detail our chemical evolution models and parameter selection in Section \ref{sec:methods}. We compare our multi-zone model predictions to the data in Section \ref{sec:multizone-results}. We discuss possible extensions to our models in Section \ref{sec:discussion}, and we and summarize our conclusions in Section \ref{sec:conclusions}.

@@

We obtain a sample of \num{171635} red giant branch and red clump stars with high-quality spectra using the selection criteria listed in Table \ref{tab:sample}, which are adapted from \citet{hayden_chemical_2015}. Table \ref{tab:uncertainties} presents the median statistical uncertainty and uncertainty dispersion ($95^{\rm th} - 5^{\rm th}$ percentile difference) of the calibrated [Fe/H] and [O/Fe]\footnote{
    In this paper, we use O as the representative $\alpha$-element in our GCE models and stellar abundance data. We refer to [O/Fe] and \aFe interchangeably, though observational studies often average several $\alpha$-element abundances to calculate \aFe.
} abundances for our sample. When calculating the galactocentric radius $R_{\rm gal}$ and midplane distance $z$ of each star, we use the \citet{bailer-jones_estimating_2021} photo-geometric distance estimates from {\it Gaia} Early Data Release 3 \citep{gaia_collaboration_gaia_2016,gaia_collaboration_gaia_2021} included in the APOGEE DR17 catalog and we adopt the Galactic coordinates of the Sun $(R,z)_\odot=(8.122, 0.0208)\kpc$ \citep{gravity_collaboration_detection_2018,bennett_vertical_2019}. 
% Table \ref{tab:apogee-regions} presents the number of APOGEE stars in bins of $R_{\rm gal}$ and $|z|$; typical distance uncertainties are much smaller than the bin width \todo{(do I need this table?)}.

\subsection{Stellar Age Estimates}
\label{sec:age-estimates}

@@

Our second age catalog utilizes the [C/N]--age relation calibrated by \citet[][submitted]{roberts_cn_2025} for red giant branch (RGB) and red clump stars. The relationship relies on the mass-dependent level of mixing during first dredge-up \citep[FDU;][]{iben_stellar_1967} to map the correlation of stellar mass, and hence age, with surface chemistry. This method has the benefit of providing age estimates for luminous giants ($\log g<2.5$), which increases the sample size at larger distances from the Sun. Limitations from the efficiency of FDU mixing and the RGB age--mass relationship mean the ages are not trustworthy outside the range $1\sim10\Gyr$, but this is the age range most useful for our purposes anyway. Additional mixing effects in low-metallicity stars %later in their RGB evolution 
also prevent the relation from being applied to luminous giant and red clump stars with $\mathFeH<-0.4$; stars on the lower RGB do not suffer from this problem. The median propagated uncertainty for the [C/N]-derived ages is $\sim1\Gyr$; however, as noted by \citet[][submitted]{roberts_cn_2025}, the propagated errors underestimate the true age uncertainty, so we enhance the uncertainties by 40\% (see Table \ref{tab:uncertainties}). With this relationship, we estimate ages for \num{113464} stars across the disk, including \num{20995} in the Solar neighborhood.

@@

% We assign a final radius and midplane distance to each stellar population as described in Section \ref{sec:migration}. 
We discuss our assumptions about the nucleosynthetic yields in Section \ref{sec:yields}, the outflow prescription in Section \ref{sec:outflows}, the star formation law in Section \ref{sec:sf-law}, the gas supply equation in Section \ref{sec:sfh}, the infall parameter selection in Section \ref{sec:parameter-selection}, and the stellar migration prescription in Section \ref{sec:migration}. We do not incorporate radial gas flows between the different zones, but we discuss their potential implications in Section \ref{sec:radial-flows}. Table \ref{tab:parameters} summarizes the model parameters and their fiducial values.

@@

        M_{\rm \star,tot}   & $5.17\times 10^{10}\Msun$     & ---   & \ref{sec:sfh} & Total stellar mass of the disk \citep{licquia_improved_2015} \\

@@

\begin{table}
    \centering
    \caption{Nucleosynthetic yields at each of the yield scales (see Section \ref{sec:yields}).}
    \include{output/yields.tex}
    \label{tab:yields}
\end{table}

@@

Unlike CCSNe, SNe Ia populate a broad distribution of delay times between progenitor formation and explosion. The time-dependent SN Ia rate per unit mass of star formation is defined as
\begin{equation}
    R_{\rm Ia}(t) = 
    \begin{cases}
        \frac{N_{\rm Ia}}{M_\star}
        \frac{f_{\rm Ia}(t)}{\int_{t_{\rm Ia}}^{t_{\rm max}} f_{\rm Ia}(t') dt'}, & t \ge t_{\rm Ia} \\
        0 & t < t_{\rm Ia},
    \end{cases}
    \label{eq:dtd-function}
\end{equation}
where $t_{\rm Ia}=40\,{\rm Myr}$ is the minimum SN Ia delay time, $t_{\rm final}=13.2\,{\rm Gyr}$ is our assumed lifetime of the Galactic disk, and $f_{\rm Ia}(t)$ is the un-normalized form of the DTD. Motivated by recent results suggesting that a large fraction of long-delayed SNe Ia improves agreement with the Milky Way's high-$\alpha$ sequence \citep{palicio_analytic_2023,dubay_galactic_2024}, we adopt a wide plateau DTD of the form
\begin{equation}
    \label{eq:plateau-dtd}
    f_{\rm Ia}(t) =
    \begin{cases}
        1, & t < 1\,{\rm Gyr} \\
        (t/1\,{\rm Gyr})^{-1.1}, & t \ge 1\,{\rm Gyr}.
    \end{cases}
\end{equation}
We also explore a simple power-law DTD,
\begin{equation}
    f_{\rm Ia}^{\rm plaw}(t) = (t/1\,\rm{Gyr})^{-1.1}.
    \label{eq:powerlaw-dtd}
\end{equation}
Equation \ref{eq:powerlaw-dtd} is consistent with SN Ia surveys \citep[e.g.,][]{maoz_star_2017} and results in more prompt Fe production.
We discuss the implications of using the power-law DTD in Section \ref{sec:abundance-distributions}.

@@

Mass-loaded outflows are a useful tool for scaling the endpoint of GCE models. \citet{weinberg_equilibrium_2017} showed that in the case of exponentially declining star formation, the O abundance approaches an equilibrium at
\begin{equation}
    \label{eq:equilibrium}
    Z_{\rm O,eq} = \frac{y_{\rm O}^{\rm CC}}{1 + \eta - r - \tau_\star/\tau_{\rm SFH}},
\end{equation}
where $r=0.4$ is the instantaneous recycling parameter, $\tau_\star$ is the star formation efficiency timescale, $\tau_{\rm SFH}$ is the $e$-folding timescale of the star formation history, and $\eta\equiv \dot\Sigma_{\rm out}/\dot\Sigma_\star$ is the outflow mass-loading factor. Motivated by Equation \ref{eq:equilibrium}, we adopt a different outflow mass-loading factor at the Solar radius $\eta_\odot\equiv\eta(R=R_\odot)$ for each of the yield sets in Table \ref{tab:yields}. Models with lower yields do not achieve a steady-state abundance in time (see Figure \ref{fig:yield-outflow}); therefore, the values of $\eta_\odot$ for \yZ{1} and \yZ{2} are lower than Equation \ref{eq:equilibrium} would indicate as required to reach Solar metallicity at the present day, assuming that $Z_{\rm O,eq}\approx Z_{\rm O,\infty}$ as suggested by \citet{johnson_milky_2024} based on stellar age trends. Table \ref{tab:models} reports the value of $\eta_\odot$ in each of our models.

@@

We note that all our models are normalized to produce the same thick-to-thin-disk mass ratio of $f_{\Sigma}(R_\odot)=0.12$ \citep{bland-hawthorn_galaxy_2016} at the Solar annulus regardless of the infall parameters. The high-$\alpha$ sequence appears much less prominent in our [O/Fe] distributions in Figure \ref{fig:twoinfall-parameters} than in the data because the one-zone model outputs include only stars formed in-situ at the Solar annulus. In our multi-zone models, most of the high-$\alpha$ stars present in the Solar neighborhood have migrated from the inner Galaxy.
